---
title: "Correlation matrix oceanographic variables"
output: html_notebook
---


```{r}
library(dplyr)
library(corrplot)
library(ggplot2)
library(tidyverse)
library(MASS) # for boxcox function
library(car)
```
Define the folder paths
```{r}
Fig_output_folder <- "C:/Users/jolaya/Documents/GitHub_projects/HRI_Fish_predictor_model/results/figures"

Table_output_folder <- "C:/Users/jolaya/Documents/GitHub_projects/HRI_Fish_predictor_model/results/tables"

```

Input data with the response and depdendent variables. This data frame contains the marine predictors created in ArcGis using as Input layer: Bathymetry 10m from TNC. Spatial resolution 10 x 10 m.

```{r}
HRI_variables <- read.csv("G:/Shared drives/NSF CoPE internal/GIS_CoPE/GIS_Belize/2_model_inputs_belize/coral_reef_modeling/01_input_csv/HRI_var_10m.csv")

# change column names
HRI_variables <- HRI_variables %>%
  rename(x = Latitude,  
         y = Longitude,  
         cor = Coral_cove, 
         algae = Algae_cove,
         fish_com = Commercial,
         fish_her = Herbivorou,       
         aspect_std_curv = aspect_s_c,
         slopeslope_240  = sloslo_240,
         slopeslope = slopeslop,
         bathy = bz_bathy
  )


```

Keep only oceanographic variables variables in the data frame.
As a result of marine predictors construction some values get -9999 but they are NA.
```{r}
colum_excluir <- c("x","y","cor","algae","fish_com","fish_her")

data <- HRI_variables[, !names(HRI_variables) %in% colum_excluir] %>%
  mutate(across(where(is.numeric), ~ifelse(. == -9999.000, NA, .)))

print(data)
```

 
```{r}
# Filtrar filas con datos completos
data_complete <- na.omit(data)  
```

Then we calculate the correlation matrix
```{r}
# Calcular la matriz de correlaci贸n
cor_matrix <- cor(data_complete, use = "complete.obs", method = "pearson")

# Save the plots
png(filename = file.path(Fig_output_folder, "correlation_all_vbles.png"), width = 6, height = 6, units = "in", res = 300)

# Visualizar la matriz de correlaci贸n
cor_all <- corrplot(cor_matrix, 
         method = "number",
         type = "upper",
         diag = FALSE,
         number.cex = 0.7)

dev.off()
```


Remove the variables that have correlations <0.7
curv_pro, curv_plan, slopeslope_240 

```{r}
exclude_vbles <- c("curv_pro", "curv_plan", "slopeslope_240")

data.cor <- data_complete[, !names(data_complete) %in% exclude_vbles]
```

Then we calculate the correlation matrix again
```{r}
# Calcular la matriz de correlaci贸n
cor_matrix.2 <- cor(data.cor, use = "complete.obs", method = "pearson")

# Save the plots
png(filename = file.path(Fig_output_folder, "correlation_final.png"), width = 6, height = 6, units = "in", res = 300)

# Visualizar la matriz de correlaci贸n
corrplot(cor_matrix.2, 
         method = "number",
         type = "upper",
         diag = FALSE,
         number.cex = 0.7)
```

```{r}
# Export table to used in the model

data_model <- HRI_variables[, !names(HRI_variables) %in% exclude_vbles] 

setwd("G:/Shared drives/NSF CoPE internal/GIS_CoPE/GIS_Belize/2_model_inputs_belize/coral_reef_modeling")

# Export using base R
write.csv(data_model, file = file.path("G:/Shared drives/NSF CoPE internal/GIS_CoPE/GIS_Belize/2_model_inputs_belize/coral_reef_modeling/01_input_csv", "HRI_variables.csv"), row.names = FALSE)
```


 













